Analyzing Crime Trends in Toronto: Insights from the Police Annual Statistical Report

### 2024-09-24 paper/paper.qmd
new change

## Reported Crimes Dataset
The dataset includes multiple fields related to crime, such as:
1. **Crime Category**: Broad crime categories like "Crimes Against Property" and "Crimes Against the Person".
2. **Crime Subtype**: Specific types of crimes (e.g., "Auto Theft", "Break & Enter").
3. **Division**: The Toronto Police divisions where crimes occurred.
4. **Incidents Reported**: The total number of reported incidents.
5. **Incidents Cleared**: The number of resolved cases.

@@

# Results

# Results

Our results are summarized in @tbl-modelresults.

@@

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

>>>>

## Reported Crimes Dataset
The dataset includes multiple fields related to crime, such as:  
**Crime Category**: Broad crime categories like "Crimes Against Property" and "Crimes Against the Person".  
**Crime Subtype**: Specific types of crimes (e.g., "Auto Theft", "Break & Enter").  
**Division**: The Toronto Police divisions where crimes occurred.  
**Incidents Reported**: The total number of reported incidents.  
**Incidents Cleared**: The number of resolved cases.

@@

# Results
## Crime Distribution by Year
The dataset reveals fluctuating crime rates between 2014 and 2023. Property crimes such as auto theft have shown consistent increases, while personal crimes, such as sexual violations, remain relatively stable.

@@

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv")) # Load the cleaned analysis data
crime_trends_plot <- analysis_data %>%
  group_by(year, crime_category) %>%
  summarise(total_incidents = sum(incidents_reported, na.rm = TRUE))  # Summing incidents per year and crime category
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-Crimes_by_Year
#| fig-cap: Crimes Reported by Year

## Create a line chart showing crimes reported by year and crime category
crime_trends_plot %>%
  ggplot(aes(x = year, y = total_incidents, color = crime_category, group = crime_category)) +
  geom_line(size = 1) +  # Line plot for crime trends
  geom_point(size = 2) +  # Adding points to lines
  labs(
    title = "Crimes Reported by Year in Toronto (2014-2023)",
    x = "Year",
    y = "Total Incidents Reported",
    color = "Crime Category"
  ) +
  theme_minimal()   # Minimal theme for clean visuals

```

According to @fig-Crimes_by_Year,
